\subsection{Completion-Based Capacity Verification}
\label{sec:verification}

The incentive-compatibility proof in \Cref{prop:bne} assumes a detection mechanism
capable of identifying over-reporting sinks within a bounded time window~$T_d$.
This section formalises the \emph{CompletionTracker} protocol that realises such
detection on-chain without external oracles.

% ─────────────────── Receipt Format ───────────────────
\subsubsection{Dual-Signed Completion Receipts}

Every task completed in BPE produces an \emph{EIP-712}~\cite{eip712} typed-data
receipt signed by both the executing sink and the requesting source:

\begin{definition}[CompletionReceipt]
\label{def:completion-receipt}
A completion receipt is a tuple
\[
  \mathcal{R} = \bigl(\tasktype,\; \texttt{sink},\; \texttt{source},\;
                       \texttt{taskId},\; t\bigr)
\]
where $\tasktype$ identifies the task type, $\texttt{sink}$ and
$\texttt{source}$ are Ethereum addresses, $\texttt{taskId}$ is a unique
32-byte identifier, and $t$ is the block timestamp at submission.
The receipt is valid iff both
$\mathrm{ECDSA.recover}(H_{712}(\mathcal{R})) = \texttt{sink}$ and
$\mathsf{msg.sender} = \texttt{source}$.
\end{definition}

Dual signing prevents unilateral fabrication: a malicious sink cannot
inflate its completions without cooperation from a source that actually
consumed the service, and a malicious source cannot credit a sink that
did not perform.

% ─────────────────── Epoch Accounting ───────────────────
\subsubsection{Rolling-Window Epoch Accounting}

Time is partitioned into epochs of duration $T_{\mathrm{epoch}} = 300$\,s.
For each (task-type, sink) pair, the contract maintains:

\begin{equation}
\label{eq:completion-rate}
r_k^{(e)} = \frac{\kappa_k^{(e)}}{\Csmooth(\tasktype, k)}
\end{equation}

\noindent where $\kappa_k^{(e)}$ is the number of valid receipts submitted
during epoch~$e$ and $\Csmooth(\tasktype, k)$ is the EWMA-smoothed declared
capacity from \Cref{eq:ewma}.  We cap $r_k^{(e)}$ at~$1$ (100\% utilisation).

When an epoch elapses, any participant may call \texttt{advanceEpoch} to
finalise the rate and reset the counter.  The on-chain cost is
${\sim}48{,}000$\,gas per epoch advance (see \Cref{sec:evaluation},
Experiment~E6).

% ─────────────────── Statistical Detection ───────────────────
\subsubsection{Statistical Divergence Detection}
\label{sec:divergence}

\begin{definition}[Under-performance event]
\label{def:underperform}
Sink~$k$ \emph{under-performs} in epoch~$e$ if
$r_k^{(e)} < \tau$, where $\tau = 0.5$ (the slash threshold in basis points,
$\mathtt{SLASH\_THRESHOLD\_BPS} = 5000$).
\end{definition}

The protocol tracks a consecutive counter~$n_k$ that increments when
$r_k^{(e)} < \tau$ and resets to zero otherwise.

\begin{proposition}[Detection Latency]
\label{prop:detection-latency}
An over-reporting sink that sustains capacity inflation
$\hat{C}_k - C_k = \varepsilon > 0$ is detected and slashed within
\begin{equation}
\label{eq:detection-time}
T_d = n_{\mathrm{thresh}} \times T_{\mathrm{epoch}} = 3 \times 300\,\text{s}
     = 900\,\text{s}
\end{equation}
provided the excess demand $\varepsilon$ induces a utilisation deficit
$r_k < \tau$ for $n_{\mathrm{thresh}} = 3$ consecutive epochs.
\end{proposition}

\begin{proof}
When $\hat{C}_k > C_k$, the additional flow routed to sink~$k$ is
$\Delta F \propto \alpha \varepsilon / \sum_j \Csmooth(j)$.  Tasks arriving
at rate exceeding true capacity~$C_k$ cannot be completed, so
$\kappa_k^{(e)} \leq C_k < \hat{C}_k$.  Hence
$r_k^{(e)} = \kappa_k / \hat{C}_k \leq C_k / (C_k + \varepsilon) < 1$.
For $\varepsilon$ large enough such that $C_k / (C_k + \varepsilon) < \tau$,
the under-performance counter increments each epoch.  After
$n_{\mathrm{thresh}} = 3$ such epochs, the auto-slash fires, proving the
claim.
\end{proof}

% ─────────────────── Auto-Slash ───────────────────
\subsubsection{Auto-Slash Mechanism}

When $n_k \geq n_{\mathrm{thresh}}$, the \texttt{advanceEpoch} function
atomically invokes \texttt{StakeManager.slash}:

\begin{equation}
\label{eq:slash-amount}
\Delta S_{\mathrm{slash}} = \frac{s_{\mathrm{bps}}}{10\,000} \cdot S_k
\end{equation}

\noindent with $s_{\mathrm{bps}} = 1000$ (10\% of stake) and a reason hash
\texttt{keccak256("COMPLETION\_UNDERPERFORMANCE")}.
After slashing:

\begin{enumerate}
  \item The consecutive counter resets ($n_k \gets 0$), giving the sink a
        recovery window.
  \item The reduced stake lowers the concave capacity cap
        $\text{cap}(S_k) = \sqrt{S_k / u}$ from \Cref{eq:capacity-cap},
        forcing the sink to either re-stake or operate at genuinely lower
        capacity---both outcomes align incentives.
  \item The epoch counter and completion tally reset for the next measurement
        window.
\end{enumerate}

\paragraph{Composability with BNE condition.}
The slash amount~\eqref{eq:slash-amount} feeds directly into
condition~\eqref{eq:bne-condition}: substituting $s = s_{\mathrm{bps}} / 10\,000$,
$T_d = 900$\,s, the BNE holds for all deviations
$\varepsilon < 0.1 \cdot S_k / (p \cdot 900)$, consistent with the
parameterisation of \Cref{sec:security}.

% ─────────────────── Future Hardening ───────────────────
\subsubsection{Future Verification Hardening}
\label{sec:future-verification}

The statistical detection mechanism of \Cref{sec:divergence} operates purely
on completion counts, requiring no access to the task payload or execution
environment.  Two complementary approaches can further reduce the trust
assumption on source--sink collusion:

\paragraph{Trusted Execution Environments (TEEs).}
Sinks running inside a TEE (e.g., Intel SGX, ARM TrustZone) can produce
hardware-attested proofs of execution.  An on-chain verifier interface
\texttt{ICapacityVerifier.verify(attestation)} can augment the
\texttt{CompletionReceipt} with a TEE quote, reducing the receipt to a
\emph{single-party} proof of work that does not require source co-signing.

\paragraph{ZK-ML proofs.}
For deterministic ML inference tasks, recent advances in zero-knowledge
machine learning (zkML)~\cite{kang2022zkml} allow sinks to generate succinct
proofs that a specific model produced a specific output.  Integrating ZK-ML
verification would upgrade the completion tracking from \emph{statistical}
(probabilistic detection over epochs) to \emph{cryptographic} (per-task
deterministic verification), at the cost of higher sink-side computation.

Both extensions are backward-compatible: the \texttt{CompletionTracker}
contract can be wrapped by a verifier contract that gates
\texttt{recordCompletion} on additional cryptographic evidence, without
modifying the epoch accounting or auto-slash logic.
